Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation.}

### 2020-07-18 journal paper/content/related_work.tex
Changed secton label in AppA, and layout/labels in AppB. Edits elsewhere for consistency.
fix broken ref

\xhdr{Bandits and exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information. MAB problems have been studied for many decades \asedit{by researchers from computer science, operations research, statistics and economics. The literature on MAB is extremely vast and multi-threaded. The most relevant thread studies a basic model of MAB with IID rewards and no auxiliary structure, see Appendix~\ref{sec:related-classes}. The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.} This literature is covered in many books and surveys. Particularly, see
\citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory. A discussion of industrial applications of MAB can be found in \citep{DS-arxiv}.

@@

\xhdr{Other related work in economics.} The competition vs. innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{Aghion-QJE05,Vives-08}. Production costs aside, this literature treats innovation as a priori beneficial for the firm. Our setting is very different, as innovation in exploration algorithms may potentially hurt the firm.

>>>>

\xhdr{Bandits and exploration.} Multi-armed bandits (\emph{MAB}) is an elegant and tractable abstraction for tradeoff between \emph{exploration} and \emph{exploitation}: essentially, between acquisition and usage of information. MAB problems have been studied for many decades \asedit{by researchers from computer science, operations research, statistics and economics. The literature on MAB is extremely vast and multi-threaded. The most relevant thread studies a basic model of MAB with IID rewards and no auxiliary structure, see Appendix~\ref{app:bg}. The basic model has been extended in many different directions, with a considerable amount of work on each: \eg payoffs with a specific structure (\eg combinatorial, linear, convex or Lipschitz), payoff distributions that change over time, and auxiliary payoff-relevant signals.} This literature is covered in many books and surveys. Particularly, see
\citet{Bubeck-survey12,slivkins-MABbook,LS19bandit-book} for background on regret-minimizing formulations, \citet{Gittins-book11} for Bayesian and Markovian formulations, and \citet{CesaBL-book,slivkins-MABbook} for connections to economics and game theory. A discussion of industrial applications of MAB can be found in \citep{DS-arxiv}.

@@

\xhdr{Other related work in economics.} The competition vs. innovation relationship and the inverted-U shape thereof have been introduced in a classic book \citep{Schumpeter-42}, and remained an important theme in the literature ever since \cite[\eg][]{aghion2005competition,Vives-08}. Production costs aside, this literature treats innovation as a priori beneficial for the firm. Our setting is very different, as innovation in exploration algorithms may potentially hurt the firm.

### 2020-07-18 journal paper/content/sec-intro.tex
fix broken ref

\xhdr{Economic interpretation: the inverted-U relationship.}
Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation, as in Figure~\ref{fig:inverted-U}.%
\footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm. It is not salient whether similar ideas and/or technologies already exist elsewhere. Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{Aghion-QJE05,Vives-08}.

>>>>

\xhdr{Economic interpretation: the inverted-U relationship.}
Interpreting the adoption of better algorithms as ``innovation", our findings can be framed in terms of the inverted-U relationship between competition and innovation, as in Figure~\ref{fig:inverted-U}.%
\footnote{Here, ``innovation" refers to adoption of a better technology at a substantial R\&D expense to a given firm. It is not salient whether similar ideas and/or technologies already exist elsewhere. Adoption of exploration algorithms tends to require substantial R\&D effort in practice, even if the algorithms themselves are well-known in the research literature \citep[\eg see][]{DS-arxiv}.}
This is a well-established concept in the economics literature, dating back to \cite{Schumpeter-42}, whereby too little or too much competition is bad for innovation, but intermediate levels of competition tend to be better \citep[\eg][]{aghion2005competition,Vives-08}.
